---
title: "Learning_shiny1"
author: "wj107"
date: "April 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learing shiny.  Part 1

That is, I assume that I'll have multiple parts to this...

So last time I created a map of Illinois, with a given county highlighted.  I was talking about creating a Shiny app so that the map could be interactive -- for example, click a button and highlight a new county.  Or a drop down menu.  Etc.

```{r IL_map, echo=FALSE}

#---required!!!!
require(ggplot2,quietly=T)
require(maps,quietly=T)

#---get USA county-level data
dat<-map_data("county")

#---focus on just illinois
IL<-subset(dat,region=="illinois")

#---find all county names
county.names<-unique(IL$subregion)
#---pick one at random
random.county<-sample(county.names,1)

#---create identifier vector... simple logical test
county.identifier<-IL$subregion==random.county
#---add to data frame
IL<-data.frame(IL,county.identifier)


#---create map, using identifier to distinguish color
p<-
#---ggplot!!
ggplot(
    #---Illinois county boundaries
    data=IL,
    #---aesthetics
    aes(
        #---long and lat
        x=long,
        y=lat,
        #---group county boundaries together
        group=subregion
    )    
)+

#---add geometry -- polygons!!
geom_polygon(
    #---set line color
    color="black",
    #---FILL COLOR DISTINGUISHED BY COUNTY IDENTIFIER!!!
    aes(fill=county.identifier)
)+

#---scale for fill allows custom setting of colors
scale_fill_manual(
    #---values is a named vector -- level and color.
    values=(
        c(
        "TRUE"="red",
        "FALSE"="grey79"
        )
    )
)+

#---no guide, err, legend, for fill
guides(fill="none")+

#---aspect ratio!!!
coord_fixed()+

#---get rid of axis titles!
xlab(NULL)+
ylab(NULL)+

#---get rid of numbers on the axes!!
scale_x_continuous(breaks=NULL)+
scale_y_continuous(breaks=NULL)


#---OUTPUT!!!
print("Last time...")
p

```

(Note while running that last code chunk, I learned a new argument to the `require` function: `quietly=T`.  Switching to R markdown forces me to learn some new tricks!)

## What is a Shiny App??

I'm still learning -- I've created a grand total of three Shiny Apps in my life.  But, at the most basic, I'd describe them saying:

- Shiny Apps create a web page
- There a various user inputs: sliders, menus, text input... I'm sure there are other input options.
- R objects -- most commonly, in my experience, graphs -- are outputed to the web page.

The input/output description leads me to think of Shiny Apps like R functions, except with an easy-to-use web interface.  In other words, I can collect the data, write some R code for analysis, but the Shiny App is key to communicating out to the world beyond folks fluent in R.  But, that said, a lot of R is needed to get a Shiny App up and running.  Which leads to me wondering...

## What does a Shiny App look like on the back end??

Reminder:  everything I'm writing about this presented with a somewhat tentative voice.  I can guess at the answer to the above question based on the three Shiny Apps I've ever made (here goes for #4) but clarifications or corrections welcome.  Just try not to present them with a "Well, actually..."  voice.  ;-)

To create a Shiny App...

- Have the `shiny` package installed and loaded
- Define a `ui` object (?) that describes the 'stuff' on the webpage -- the user inputs, the R object output
- Define a `server` function that does the R coding in the background
- Call the function `shinyApp(ui, server)` to put it all together into a web page

Let's make a simple Shiny App example:  User inputs a number, R generates a random normal distribution of N values, output is a histogram.

So the `ui` object needs to get an input, and display a graph.

```{r create_ui}
#---required!!!
require(shiny, quietly=T)

#---what's on the web page???
ui<-
  #---yeah I don't know about this.  Straight copying the help page, now
  fluidPage(
    #####INPUT
    #---numeric input!!
    numericInput(
      #---what variable does R call the input vale??
      "N",
      #---a prompt for the user
      "How many observations for a random normal distribution??",
      #---initial value!!
      25
    ),
    ####OUTPUT
    #---plot a graph!
    plotOutput(
      #---name of the plot object
      "histogram"
    )
    
  )
  
```


My first thoughts after typing that out:

- I have no idea about `fluidPage`, I copied that from the help page.  But, in time, I'll learn a few more arguments for that.
- There's two main pieces for the `fluidPage` function.  An input part, and an output part.  Makes sense -- an interactive webpage needs both.
- The variable names: `N` and `histogram`, are in quotes.
- For the `plotOutput` part, I'm thinking ahead -- I haven't created ANY output yet.  But, I will, and when I do it'll be a plot named `histogram`.

(again, if you're curious, I'm writing all this without knowing exactly what I'm going to get.  I have an idea, I've made other Shiny Apps... but trying to talk my way through all the details as I re-create a new one, here)

So, from the list above I need a `server` function that tells R what to do.  And, as mentioned above, I want it to get a random normal distribution, and create a histogram called `histogram`.

```{r create_server}

#---what's R doing in the background??
server<-
  #---best I can tell, this is ALWAYS the next line.  You're making a function with inputs and outputs.
  function(input, output){
    #---create the output histogram...
    output$histogram<-
      #---renderPlot...not 100% sure about this.
      renderPlot(
        #---create a histogram...
        hist(
          #---...of random normal data
          rnorm(
            #---with N values, where N is the input
            input$N
          )
        )
      )
    
  }

```

Hot takes after writing that:

- The first two lines are boilerplate... EVERY `server` function for a shiny app starts like that.
- Calling the input/output variables or objects from `ui`:  you don't use the name alone, or in quotes (like in `ui`).  It's `input$xxx` or `output$xxx`.
- I think, to the best of my understanding, `renderPlot` is something to tell the `server` function that the stuff inside might change based on the web page input.  So, if there's fixed stuff -- like if you want the normal distribution to have `mean=10`, you could add that code in somewhere else.  But, if it changes -- like the `input` stuff, natch -- it needs to be inside a `renderPlot` function.  I think there's other similar functions to indicate 'changing' pieces within `server`.  I don't know them... yet.

There's an `ui` object.  There's a `server` function.  Time to see if it works -- time to call `shinyApp`!


```{r create_shiny, eval=FALSE}
shinyApp(ui,server)
```

Ha -- I forgot how to prevent a chunk from running, `eval=FALSE` (thanks Google!), I used that about a year ago writing my [Master's Thesis](https://github.com/wj107/495project) in LaTeX with R chunks (I believe that's called an `Rnw` file?  How soon I forget...) 

So, Markdown being static, and Shiny App's being dynamic, there's no way to see if this works without copying the code from the last three chunks into an R script, and running it.  Will it work?  I don't know, honestly.  You should try it too!  Here goes...