---
title: "L5-maps4"
author: "wj107"
date: "April 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Maps, part 4

Per usual, last time I had grand goals, but only made it about two or three steps in.  Whatever.  I was grinding, figuring out some stuff... and made a decently cool map at the end, form KML data found online!  

```{r chicago_base_map, echo=FALSE}

#---create precinct boundary R object
precinct.boundaries<-
    #----load KML data
    getKMLcoordinates(
        #---load my desired KML file
        "Chicago_Precincts.kml",
        #---ignore altitude... I'm dealing with 2D data!!
        ignoreAltitude=TRUE
    )

#---reassign precinct boundary
precinct.boundaries<-
    #---for each of the polygons in the boundary data...
    lapply(precinct.boundaries,
          #---...convert it to its own data frame
           as.data.frame
          )

#---reassign precinct boundary data
precinct.boundaries<-
    #---bind each of the polygons into one big data frame
    bind_rows(precinct.boundaries,
            #---distinguish each polygon!
            .id="polygon"
    )

#---rename columns...
names(precinct.boundaries)<-c("precinct","long","lat")



```

I mentioned my next goal:  get a cool background to the map, using `ggmap`.  But, since Google required registration to use `ggmap`, I needed to shift to my desktop computer (I was on a chromebook) to get situated with `ggmap`.  But, in the process, I realized that my desktop wasn't ready to mess with KML data right away.  I needed to get some extra packages installed.

First, `rdgal`.  But, I kept getting an error message: `configure: error: gdal-config not found or not executable.`  Hmmm.  [Stack overflow tells me that it's because there's *development* packages for geospatial data in Linux that I'm missing.](https://stackoverflow.com/a/12143411)  (My desktop computer runs on Linux)  Ran `apt-cache search libgdal` in `bash`, found and installed the packages, and got `rgdal` running in R.  Hat tip to [Dirk EddelBuettel](https://stackoverflow.com/users/143305/dirk-eddelbuettel) for the advice.  You know, I've come across his name various times looking for R help.  I get the sense he knows what's up.

